---
title: "Use BreedSimCost"
author: "Jean-Luc Jannink"
date: "5/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages  
```{r load packages, message=FALSE}
ip <- installed.packages()
packages_used <- c("AlphaSimR", "tidyverse", 
                   "workflowr", "here", "devtools")
for (package in packages_used){
  if (!(package %in% ip[,"Package"])) install.packages(package)
}

packages_devel <- c("BreedSimCost")
for (package in packages_devel){
  if (!(package %in% ip[,"Package"])){
    devtools::install_github(paste0("jeanlucj/", package), ref="main",
                             build_vignettes=F)
  }
}
library(BreedSimCost)

here::i_am("analysis/BreedSimCostTest.Rmd")

random_seed <- 45678
set.seed(random_seed)
```

### Illustrate using the functions in BreedSimCost  

#### Initialize the breeding program and optimization  
Initializing the optimization relies on four text files.  
1. File with parameters that control the population history and genetic 
architecture of founders  
2. File with parameters that control the breeding scheme  
3. File with parameters that control the costs of different breeding resources  
4. File with parameters that control how optimization is achieved  
```{r Initialize program}
bsd <- initializeProgram(
         here::here("data", "FounderCtrlFile.txt"),
         here::here("data", "SchemeCtrlFile.txt"),
         here::here("data", "CostsCtrlFile.txt"), 
         here::here("data", "OptimizationCtrlFile.txt")
                         )
```

#### Fill the variety development pipeline  
Initially, the VDP is empty.  Here are commands that simulate a number of years
of phenotypic selection that might precede the initiation of a two-part strategy  
```{r Fill variety development pipeline}
# Year 1
bsd$year <- bsd$year+1
bsd <- makeVarietyCandidates(bsd)

entriesSDN <- bsd$varietyCandidates@id
bsd <- runVDPtrial(bsd, "SDN", entriesSDN)

parents <- selectParentsBurnIn(bsd)
bsd <- makeCrossesBurnIn(bsd, parents)

# Year 2
bsd$year <- bsd$year+1
bsd <- makeVarietyCandidates(bsd)

entries <- chooseTrialEntries(bsd, toTrial="SDN")
bsd <- runVDPtrial(bsd, "SDN", entries)
entries <- chooseTrialEntries(bsd, fromTrial="SDN", toTrial="CET")
bsd <- runVDPtrial(bsd, "CET", entries)

parents <- selectParentsBurnIn(bsd)
bsd <- makeCrossesBurnIn(bsd, parents)

# Year 3 and onward
for (burnIn in 1:bsd$nBurnInCycles){
  bsd$year <- bsd$year+1
  bsd <- makeVarietyCandidates(bsd)
  
  entries <- chooseTrialEntries(bsd, toTrial="SDN")
  bsd <- runVDPtrial(bsd, "SDN", entries)
  entries <- chooseTrialEntries(bsd, fromTrial="SDN", toTrial="CET")
  bsd <- runVDPtrial(bsd, "CET", entries)
  entries <- chooseTrialEntries(bsd, fromTrial="CET", toTrial="PYT")
  bsd <- runVDPtrial(bsd, "PYT", entries)

  parents <- selectParentsBurnIn(bsd)
  bsd <- makeCrossesBurnIn(bsd, parents)
}
plot(AlphaSimR::gv(bsd$varietyCandidates))

burnedInBSD <- bsd
```

#### Code to simulate one realization of the breeding scheme  
This code is not evaluated in the Rmarkdown but is shown here to illustrate
how you would write this function so that it embodies different schemes  
The scheme you want to run has to be scripted into the function `runWithBudget`  
```{r runWithBudget, eval=FALSE}
# This is code from the package that you can run
# In the package, it's in the function runWithBudget
  startValues <- calcCurrentStatus(bsd)
  
  for (twoPart in 1:bsd$nCyclesToRun){
    bsd$year <- bsd$year+1
    bsd <- makeVarietyCandidates(bsd)
    
    # Within the year, sequentially run the VDP...
    for (stage in 1:bsd$nStages){
      entries <- chooseTrialEntries(bsd, toTrial=bsd$stageNames[stage],
                   fromTrial=ifelse(stage == 1, NULL, bsd$stageNames[stage-1]))
      bsd <- runVDPtrial(bsd, bsd$stageNames[stage], entries)
    }

    # ... And then run the PIC. Sequencing could be made more complicated
    for (genSel in 1:bsd$nPopImpCycPerYear){
      optCont <- selectParents(bsd)
      bsd <- makeCrosses(bsd, optCont)
    }
  }
  
  endValues <- calcCurrentStatus(bsd)

  print(startValues)
  print(endValues)
```

#### The main interface for optimizing budgets
`optimizeByLOESS` is the function that does all of the optimization tasks.  
```{r Cost optimization, message=FALSE}
bsd <- burnedInBSD
optimizationResults <- optimizeByLOESS(bsd)
saveRDS(optimizationResults, here::here("data", "_optRes.rds"))
```

#### Plotting function
`plotLoessPred` makes the hexbin plot showing responses at different budgets. 
The red circle shows the budget found to be optimal.  
Also plotted is one iteration of the budget found to be optimal  
```{r Plot results, results='hold'}
plp <- plotLoessPred(resultMat=optimizationResults$results)
bestPerc <- plp$percMean
cat(" Budget allocations with highest gain\n", round(unlist(bestPerc), 3), "\n")
bsd <- burnedInBSD
burnInInd <- AlphaSimR::nInd(bsd$varietyCandidates)
bsd <- runWithBudget(bestPerc, bsd, returnBSD=T)
rangeGV <- range(AlphaSimR::gv(bsd$varietyCandidates))
plot(AlphaSimR::gv(bsd$varietyCandidates))
lines(c(burnInInd, burnInInd), rangeGV, lwd=3, col="red")
```

